---
title: "DATA 555 Final Project"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

# Main Dashboard


```{r setup, include=FALSE}
library(ggplot2)
library(plotly)
library(dplyr)
library(DT)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## **Side by Side Boxplot: Age Distribution by Relapse Status (Plotly)**

```{r boxplot-relapse, echo=FALSE}
data <- read.csv("/Users/annachen/Desktop/semester 4/DATA 555/sex_age_data.csv")
data$relapse <- factor(data$relapse, labels = c("No Relapse", "Relapse"))
data$sex <- factor(data$sex, labels = c("Female", "Male"))

p1 <- ggplot(data, aes(x = relapse, y = age, fill = sex)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +
  geom_jitter(aes(text = paste("Age:", age, "<br>Sex:", sex, "<br>Treatment Duration:", duration)),
              width = 0.2, alpha = 0.7) +
  labs(
    title = "Age Distribution by Relapse Status",
    x = "Relapse Status of Cancer",
    y = "Age (years)",
    fill = "Sex"
  ) +
  theme_minimal()

ggplotly(p1, tooltip = "text")
```

### **Key Takeaways**
- **Older patients tend to relapse more frequently**.
- The **distribution of male and female patients within relapse groups** is clearly distinguishable.
- **Jitter points show individual patient details** when hovered over.

### **Real-World Impact and Importance of Dashboard**

- Understanding relapse patterns by age and gender can inform more personalized follow-up treatment plans.
- This tool enables interactive exploration of patient subgroups to support data-driven healthcare decisions.


## **Scatter Plot: Age vs. Treatment Duration (Plotly)**

```{r scatter-age-duration, echo=FALSE}
p2 <- ggplot(data, aes(x = age, y = duration, color = factor(relapse))) +
  geom_point(aes(text = paste("Age:", age, "<br>Duration:", duration, "<br>Relapse Status of Cancer:", relapse)),
             alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
  labs(
    title = "Scatter Plot: Age vs. Treatment Duration",
    x = "Age (years)",
    y = "Treatment Duration (years)",
    color = "Relapse Status of Cancer"
  ) +
  theme_minimal()

ggplotly(p2, tooltip = "text")
```

### **Key Takeaways**
- **No strong correlation observed** between age and treatment duration.
- Patients who **relapsed tend to have slightly longer treatment durations**.
- **Hover tooltips display patient-specific information**.


### **View Source Code on GitHub**

https://github.com/Anna-YuanChen/data555-final-project/









# Dataset Description

The dataset comprises clinical data from 90 patients, containing demographic and treatment-related variables to investigate how age, sex, and treatment duration influence relapse outcomes.

- Data Source: Patient records with demographic and treatment-related information. Collecting from the BIOS 509 course material.
- Sample Size: 90 individuals.
- Data Collection Method: Extracted from clinical observations, recording age, sex, treatment duration, and relapse status.
- Timeframe & Population: Data reflects patients undergoing treatment, with relapse status recorded as a key outcome.









